---
title: "Fitting A Logistic Regression with Aggregate Data in R"
subtitle: "Example Using Simulated Migration Data"
author: "Francisco Rowe"
date: "`r Sys.Date()`"
output: tint::tintHtml
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```

# Aim

This notebook illustrates three ways to estimate a logistic regression model based on individual-level data and aggregate data and arrive to the same model estimates. 

# Dependencies

```{r}
library(tidyverse)
```


# Data

For the purposes of this notebook, let's assume we are interested in fitting a migration model and we start with individual level data. We assume that the decision to migration is a function of `gender` and `age`. The decision to migrate is captured with a binary variable: 1 if an individual migrated in the last year; 0 otherwise. Here we use simulated data so let's hope that our estimates will be representative of a real migration process.

```{r}
rm(list=ls())
set.seed(5)
df <- tibble(gender = as.factor(sample(c("male","female"), 
                                       300, 
                                       replace = TRUE, 
                                       prob=c(0.6, 0.4))),
           age_bracket = as.factor(sample(c("[15<25]","[25-45]", "[45-65]"), 
                                          300, 
                                          replace = TRUE, 
                                          prob=c(0.35, 0.55, 0.1))),
           migrate = rbinom(300, 
                             1, 
                             prob = 0.2))

df
```


# Logistic Regression on Individual-Level Data

Let's first a logistic regression on individual level data

```{r}
eq1 <- migrate ~ gender + age_bracket
m1 <- glm(eq1, data = df, family = binomial("logit"))
summary(m1)
```

# Logistic Regression on Aggregate Data

Now let's fit the same model on aggregate data. To this end, we first need to aggregate the data:

```{r}
df_agg<-df %>% group_by(gender, age_bracket) %>%
  summarise( pop_count = n(), 
             moves = sum(migrate)) %>%
  ungroup() %>% 
  mutate(prob_mig = moves / pop_count)

df_agg
```

Let's now estimate the model:

```{r}
eq2 <- prob_mig ~ gender + age_bracket
m2 <- glm(eq2,
          data=df_agg,
          weights = pop_count, 
          family = binomial("logit"))
summary(m2)
```

Compare these estimates to our individual-level model and note that we arrive to the same model estimates!

# Logistic Regression Using cbind

Let's try a second alternative and use `cbind`. It is useful to know this if you ever want to fit a multinomial logistic model in R using the package [`VGAM`](https://cran.r-project.org/web/packages/VGAM/index.html)

```{r}
df_agg$stayers <- df_agg$pop_count - df_agg$moves

eq3 <- cbind(moves, stayers) ~ gender + age_bracket

m3 <- glm(eq3, 
          data = df_agg,
          family = binomial("logit"))
summary(m3)
```

For more on logistic regression, check out this course: [Introduction to Statistical Learning in R](https://fcorowe.github.io/sl/)
